---
title: "SSA_Mexico_Data"
author: "Jose Tamez"
date: "4/19/2020"
output: html_document
---


```{r, echo = FALSE}
#knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# Covid19 Survival

This document does the analysis of the patient level data set.


Loading the libraries

```{r results = "hide" }
library("FRESA.CAD")
library(epiR)
library(fastAdaboost)
repetitions = 300
trainFraction = 0.90

```

## Load the data

```{r}

#COVID19MEXICO <- read.csv("./200420COVID19MEXICO.csv",
#                           na.strings="99",stringsAsFactors = FALSE)
COVID19MEXICO <- read.csv("C:/Users/eider/Dropbox/AD experiments/Covid_mexico/200503COVID19MEXICO.csv",na.strings="99",stringsAsFactors = FALSE)
#COVID19MEXICO <- read_csv("200503COVID19MEXICO.csv",na = "99")

```

## Cleaning the data
 se seleccionan las variables para crear los modelos
 se seleccionan solo los casos positivos de covid
 se seleccionan los que tienen fecha de defuncion valida o entraron a cuidados intensivos como outome =1 , los demas 0
 

```{r}
colnamesMexico <- colnames(COVID19MEXICO)

varnames <- c("SEXO","TIPO_PACIENTE","FECHA_INGRESO","FECHA_SINTOMAS","FECHA_DEF",
              "INTUBADO","NEUMONIA","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR",
              "HIPERTENSION","OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA",
              "TABAQUISMO","OTRO_CASO","RESULTADO","UCI")

varnames[!(varnames %in% colnamesMexico)]
colnamesMexicoComplete <- COVID19MEXICO[,varnames]
colnamesMexicoComplete <- colnamesMexicoComplete[complete.cases(colnamesMexicoComplete),]

for (vn in varnames) {
  print(vn)
  tb <- table(colnamesMexicoComplete[,vn])
  if (length(tb)<5){
    colnamesMexicoComplete[,vn] <- 1*(colnamesMexicoComplete[,vn] == 1)
    tb <- table(colnamesMexicoComplete[,vn])
  }
  print(tb)
}

varforDiagnosisCovid <- c("SEXO","FECHA_DEF","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR","HIPERTENSION","OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA","TABAQUISMO","RESULTADO","UCI")
#selecting just the variables needed 
datosDiagnosticoAlta <- colnamesMexicoComplete[,varforDiagnosisCovid] 
#solo los que dieron positivo a covid
datosDiagnosticoAlta <- subset(datosDiagnosticoAlta, RESULTADO == 1)
datosDiagnosticoAlta$RESULTADO <- NULL
#pacientes que no tienen una fecha insespecificada de defuncion O si entraton a cuidados intensivos
datosDiagnosticoAlta$outcome <- 1*(datosDiagnosticoAlta$FECHA_DEF != "9999-99-99" | datosDiagnosticoAlta$UCI == 1)
datosDiagnosticoAlta$FECHA_DEF <- NULL
datosDiagnosticoAlta$UCI <- NULL

table(datosDiagnosticoAlta$outcome)

```


```{r}
datosDiagnosticoAlta.mat <- as.data.frame(model.matrix(outcome ~ SEXO*.,datosDiagnosticoAlta))

datosDiagnosticoAlta.mat$`SEXO:EDAD` <- NULL

#datosDiagnosticoAlta.mat <- as.data.frame(model.matrix(outcome ~ (SEXO+DIABETES+HIPERTENSION)*.,datosDiagnosticoAlta))
#datosDiagnosticoAlta.mat$SEXO.EDAD <- NULL
#datosDiagnosticoAlta.mat$`DIABETES:EDAD` <- NULL
#datosDiagnosticoAlta.mat$`HIPERTENSION:EDAD` <- NULL


datosDiagnosticoAlta.mat$`(Intercept)` <- NULL
datosDiagnosticoAlta.mat$outcome <- as.numeric(datosDiagnosticoAlta$outcome)
fnames <- colnames(datosDiagnosticoAlta.mat)
fnames <- str_replace_all(fnames," ","_")
fnames <- str_replace_all(fnames,"/","_")
fnames <- str_replace_all(fnames,":","_")
fnames <- str_replace_all(fnames,"//.","_")
colnames(datosDiagnosticoAlta.mat) <- fnames


datosDiagnosticoAlta <- datosDiagnosticoAlta.mat


caseSet <- subset(datosDiagnosticoAlta,outcome==1)
controlSet <- subset(datosDiagnosticoAlta,outcome==0)

# 70 % of the covid cases
samplesCasetrain <- sample(nrow(caseSet),0.70*nrow(caseSet))
# subseting the 70% 
trainingCaseset <- caseSet[samplesCasetrain,]

# 500 control cases
samplesControltrain <- sample(nrow(controlSet),500)
#subseting 500 control cases
trainingControlset <- controlSet[samplesControltrain,]

table(datosDiagnosticoAlta$outcome)
#merge 70% of data with cases and 500 training scores
trainingset <- rbind(trainingCaseset,trainingControlset)

# selecting the remaining 30% of positive covid cases
controltestset <- controlSet[-samplesControltrain,]
# merge the remaining 30 and the controlsetttest
testingset <- rbind(caseSet[-samplesCasetrain,],controltestset[sample(nrow(controltestset),1000),])

table(trainingset$outcome)
plot.new()
op <- par(no.readonly = TRUE)

```



```{r}
HLCM_COVID <- HLCM_EM(outcome ~.,trainingset )

table(HLCM_COVID$classSet,trainingset$outcome)

#unique(var$classModel)

HLCM_COVID_2 <- HLCM(outcome ~.,trainingset )
table(HLCM_COVID_2$classSet,trainingset$outcome)

```

```{r}
HLCM_COVID_3 <- HLCM( outcome ~.,trainingset, method=adaboost ,
                      asFactor = TRUE,
                  nIter=10)
table(HLCM_COVID_3$classSet,trainingset$outcome)

HLCM_COVID_4 <- HLCM_EM( outcome ~.,trainingset, method=adaboost ,
                      asFactor = TRUE,
                  nIter=10)
table(HLCM_COVID_4$classSet,trainingset$outcome)

```



```{r}
trainingset_norm <- scale(trainingset)

data_clsuter_4<- cbind(trainingset_norm,cluster=HLCM_COVID_4$classSet)
data_clsuter_4 <- as.data.frame(data_clsuter_4)


 ggplot(data = data_clsuter_4, aes(x=as.factor(SEXO), y=EDAD)) + 
             geom_boxplot(aes(fill=as.factor(SEXO)))+
 facet_wrap( ~ as.factor(cluster), scales="free") +
   ggtitle("HCLM_EM + ADABOOST CLUSTERS ") +
   theme(legend.position = "none")
 


```





```{r}

 data_clsuter_3<- cbind(trainingset_norm,cluster=HLCM_COVID_3$classSet)
data_clsuter_3 <- as.data.frame(data_clsuter_3)


 ggplot(data = data_clsuter_3, aes(x=as.factor(SEXO), y=EDAD)) + 
             geom_boxplot(aes(fill=as.factor(SEXO)))+
 facet_wrap( ~ as.factor(cluster), scales="free") +
   ggtitle("HCLM + ADABOOST CLUSTERS ") +
   theme(legend.position = "none")
 
```

```{r}


 data_clsuter_2<- cbind(trainingset_norm,cluster=HLCM_COVID_2$classSet)
data_clsuter_2 <- as.data.frame(data_clsuter_2)


 ggplot(data = data_clsuter_2, aes(x=as.factor(SEXO), y=EDAD)) + 
             geom_boxplot(aes(fill=as.factor(SEXO)))+
 facet_wrap( ~ as.factor(cluster), scales="free") +
   ggtitle("HCLM + BSWIMS CLUSTERS ") +
   theme(legend.position = "none")
 
 
#p + facet_wrap( ~ cluster, scales="free")


```


```{r}
 data_clsuter_1<- cbind(trainingset_norm,cluster=HLCM_COVID$classSet)
data_clsuter_1 <- as.data.frame(data_clsuter_1)


 ggplot(data = data_clsuter_3, aes(x=as.factor(SEXO), y=EDAD)) + 
             geom_boxplot(aes(fill=as.factor(SEXO)))+
 facet_wrap( ~ as.factor(cluster), scales="free") +
   ggtitle("HCLM_EM + BSWIMS CLUSTERS ") +
   theme(legend.position = "none")
```


```{r}
names(HLCM_COVID_4$classSet) <- rownames(trainingset)





columns <- list()
index =1
for (cluster in unique(HLCM_COVID_4$classSet)) {
  
  columns[[index]] <-  HLCM_COVID_4$classSet == cluster
  index= index+1
}
index=1
cluster_i <- list()
for (column in columns){
  cluster_i[[index]] = trainingset[column,]
  index=index+1
}


```

## logistic Model

```{r results = "hide", dpi=600, fig.height= 6.0, fig.width= 6.0}
lmodel <- glm(outcome~.,trainingset,family=binomial(link=logit))

```

### Logistic Results

```{r results = "asis", dpi=600, fig.height= 6.0, fig.width= 6.0}
par(op)

cStats <- predictionStats_binary(cbind(trainingset$outcome,predict(lmodel,trainingset)),plotname = "Logit");


sm <- summary(lmodel)
pander::pander(sm$coefficients)

testResult <-  predict(lmodel,testingset)
testTable <- table(testResult < 0,testingset$outcome==0)
pt <- epi.tests(testTable)

cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "Logit Testing");

pander::pander(cStats$ROC.analysis$ROC.AUCs)
pander::pander(pt$tab)
pander::pander(pt$rval)
```


```{r results = "hide", dpi=600, fig.height= 6.0, fig.width= 6.0}
bm <- BSWiMS.model(outcome ~ .,trainingset,NumberofRepeats = -100)

```

### Results

```{r results = "asis", dpi=600, fig.height= 6.0, fig.width= 6.0}
par(op)
pander::pander(bm$univariate)

cStats <- predictionStats_binary(cbind(trainingset$outcome,predict(bm,trainingset)),plotname = "BSWiMS");
plot(predict(bm,trainingset),predict(bm$bagging$bagged.model,trainingset))

sm <- summary(bm)
pander::pander(sm$coefficients)

par(op)

#gplots::heatmap.2(bm$bagging$formulaNetwork,trace="none",mar = c(10,10),main = "B:SWiMS Formula Network",cexRow = 0.75,cexCol = 0.75)

testResult <-  predict(bm,testingset)
testTable <- table(testResult < 0.5,testingset$outcome==0)

ptbs <- epi.tests(testTable)

cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "BSWiMS Testing");
cStats <- predictionStats_binary(cbind(testingset$outcome,predict(bm$bagging$bagged.model,testingset)),plotname = "BSWiMS Testing");
plot(predict(bm,testingset),predict(bm$bagging$bagged.model,testingset))

pander::pander(cStats$ROC.analysis$ROC.AUCs)
pander::pander(ptbs$rval)
pander::pander(ptbs$tab)




```



### Riks Charts


```{r results = "asis", dpi=600, fig.height= 6.0, fig.width= 6.0}
par(op)

colfeat <- rownames(bm$bagging$formulaNetwork)

riskhealtyage <- as.data.frame(cbind(outcome=rep(0,71),EDAD=c(20:90)))
for (cname in  colfeat[-1])
{
  riskhealtyage <- cbind(riskhealtyage,rep(0,nrow(riskhealtyage)))
}
colnames(riskhealtyage ) <- c("outcome","EDAD",colfeat[!(colfeat %in% c("EDAD"))])
baserisktable <- riskhealtyage
colfeat <- colnames(riskhealtyage)[-1]
baseprobability <- predict(bm,riskhealtyage);
plot(baseprobability~riskhealtyage$EDAD,type="l",col=1,xlab="Age",ylab="p(Death or ICU)",ylim=c(0,1),main="Probability of Death or ICU (Postive COVID-19)",lty=1,lwd=2)

nc=2;
for (cname in  colfeat[!(colfeat %in% c("EDAD"))])
{
  nriskhealtyage <- riskhealtyage; 
  nriskhealtyage[,cname] <- rep(1,nrow(riskhealtyage)) 
   riskprobability <- predict(bm,nriskhealtyage);
  lines(riskprobability~riskhealtyage$EDAD,col=nc,lty=nc,lwd=2)
   nc <- nc + 1
}

leynames <- c("Baseline",colfeat[-1])  
legend("topleft",legend = leynames,lty=c(1:length(colfeat)),col =c(1:length(colfeat)),lwd=2,cex=0.75)

plot(baseprobability~riskhealtyage$EDAD,type="l",col=1,xlab="Age",ylab="p(Death or ICU)",ylim=c(0,1),main="Probability of Death or ICU Diabetes (Postive COVID-19)",lty=1,lwd=2)
riskhealtyage[,"DIABETES"] <- rep(1,nrow(riskhealtyage)) 
nc=2;
for (cname in  colfeat[!(colfeat %in% c("EDAD","DIABETES"))])
{
  nriskhealtyage <- riskhealtyage; 
  nriskhealtyage[,cname] <- rep(1,nrow(riskhealtyage)) 
   riskprobability <- predict(bm,nriskhealtyage);
  lines(riskprobability~riskhealtyage$EDAD,col=nc,lty=nc,lwd=2)
   nc <- nc + 1
}
legend("topleft",legend = leynames[!(leynames %in% c("DIABETES"))],lty=c(1:length(colfeat)),col =c(1:length(colfeat)),lwd=2,cex=0.75)

plot(baseprobability~riskhealtyage$EDAD,type="l",col=1,xlab="Age",ylab="p(Death or ICU)",ylim=c(0,1),main="Probability of Death or ICU Hipertension (Postive COVID-19)",lty=1,lwd=2)
riskhealtyage <- baserisktable
riskhealtyage[,"HIPERTENSION"] <- rep(1,nrow(riskhealtyage)) 
nc=2;
for (cname in  colfeat[!(colfeat %in% c("EDAD","HIPERTENSION"))])
{
  nriskhealtyage <- riskhealtyage; 
  nriskhealtyage[,cname] <- rep(1,nrow(riskhealtyage)) 
   riskprobability <- predict(bm,nriskhealtyage);
  lines(riskprobability~riskhealtyage$EDAD,col=nc,lty=nc,lwd=2)
   nc <- nc + 1
}
  
legend("topleft",legend = leynames[!(leynames %in% c("HIPERTENSION"))],lty=c(1:length(colfeat)),col =c(1:length(colfeat)),lwd=2,cex=0.75)


plot(baseprobability~riskhealtyage$EDAD,type="l",col=1,xlab="Age",ylab="p(Death or ICU)",ylim=c(0,1),main="Probability of Death or ICU Hipertension and Diabetes (Postive COVID-19)",lty=1,lwd=2)
riskhealtyage <- baserisktable
riskhealtyage[,"HIPERTENSION"] <- rep(1,nrow(riskhealtyage)) 
riskhealtyage[,"DIABETES"] <- rep(1,nrow(riskhealtyage)) 
nc=2;
for (cname in  colfeat[!(colfeat %in% c("EDAD","HIPERTENSION","DIABETES"))])
{
  nriskhealtyage <- riskhealtyage; 
  nriskhealtyage[,cname] <- rep(1,nrow(riskhealtyage)) 
   riskprobability <- predict(bm,nriskhealtyage);
  lines(riskprobability~riskhealtyage$EDAD,col=nc,lty=nc,lwd=2)
   nc <- nc + 1
}
  
legend("topleft",legend = leynames[!(leynames %in% c("HIPERTENSION","DIABETES"))],lty=c(1:length(colfeat)),col =c(1:length(colfeat)),lwd=2,cex=0.75)

```

