---
title: "covid_mexico_clusters"
author: "eider"
date: "5/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r }
library("FRESA.CAD")
library(epiR)
library(fastAdaboost)
library(fmsb)
library(colormap)
library(Rtsne)
library(randomForest)


setwd("C:/Users/eider/Dropbox/AD experiments/Covid_mexico/")
```



# Cleaning the data
 se seleccionan las variables para crear los modelos
 se seleccionan solo los casos positivos de covid
 se seleccionan los que tienen fecha de defuncion  o entraron a cuidados intensivos como outome =1 , los demas 0
 
```{r}
#COVID19MEXICO <- read.csv("200512COVID19MEXICO.csv",na.strings="99",stringsAsFactors = FALSE)

#colnamesMexico <- colnames(COVID19MEXICO)
#
#varnames <- c("SEXO","TIPO_PACIENTE","FECHA_INGRESO","FECHA_SINTOMAS","FECHA_DEF",
#              "INTUBADO","NEUMONIA","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR",
#              "HIPERTENSION","OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA",
#              "TABAQUISMO","OTRO_CASO","RESULTADO","UCI")
#
#
#colnamesMexicoComplete <- COVID19MEXICO[,varnames]
#colnamesMexicoComplete <- colnamesMexicoComplete[complete.cases(colnamesMexicoComplete),]
#
#varforDiagnosisCovid <- #c("SEXO","FECHA_DEF","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR","HIPERTENSION",#"OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA","TABAQUISMO","RESULTADO","UCI")
#
## selecting just the variables needed 
#datosDiagnosticoAlta <- colnamesMexicoComplete[,varforDiagnosisCovid] 
#
## solo los que dieron positivo a covid
#datosDiagnosticoAlta <- subset(datosDiagnosticoAlta, RESULTADO == 1)
##save(datosDiagnosticoAlta,file="datosDiagnosticoAlta.RDATA")
load("datosDiagnosticoAlta.RDATA")
```
 
 
```{r}
datosDiagnosticoAlta$RESULTADO <- NULL
#pacientes difuntos O si entraton a cuidados intensivos (UCI)
datosDiagnosticoAlta$outcome <- 1*(datosDiagnosticoAlta$FECHA_DEF != "9999-99-99" | datosDiagnosticoAlta$UCI == 1)
datosDiagnosticoAlta$outcome <-NULL
datosDiagnosticoAlta$FECHA_DEF <- NULL
datosDiagnosticoAlta$UCI <- NULL

table(datosDiagnosticoAlta$outcome)
```
 
 
#spliting the dataset 
- deberia de balancear mejor el split de train y test? 
```{r}
set.seed(42)
caseSet <- subset(datosDiagnosticoAlta,outcome==1)
controlSet <- subset(datosDiagnosticoAlta,outcome==0)

# 70 % of the UCI or death cases
samplesCasetrain <- sample(nrow(caseSet),0.70*nrow(caseSet))
# subseting the 70% 
trainingCaseset <- caseSet[samplesCasetrain,]

# 500 control cases
samplesControltrain <- sample(nrow(controlSet),500)
#subseting 500 control cases
trainingControlset <- controlSet[samplesControltrain,]
#merge 70% of data with cases and 500 training scores
trainingset <- rbind(trainingCaseset,trainingControlset)
#######################TEST

# selecting the remaining 30% of UCI and Death
controltestset <- controlSet[-samplesControltrain,]
# merge the remaining 30 and the controlsetttest
testingset <- rbind(caseSet[-samplesCasetrain,],controltestset[sample(nrow(controltestset),1000),])

```




preguntas:
-deberia de filtrar a los clusters que tengan mayor a algun treshold de observaciones?
-alguna recomendacion para el hyperparamater del TSNE?

#func para sacar las LC
```{r}
getLattentClasses <- function(lattentClassModel,testSet){
  #get the results from test set
  testResult <-  predict(lattentClassModel,testSet)
  #from each observation get the max probabilty to belong to a given LC
  lattentclass <- apply(attr(testResult,"probabilities"),1,which.max)
  cat("Lattent Classes found: ",length(unique(lattentclass)))
  lattentclass
  
}
```

 
 
#func para ver los radar plot de cada clase y su respectivo TSNE
```{r}
#data must be in feature-outcome format
radar_and_tsne_from_latentclass <- function(theData,theClasses){
  #combinamos las clases resultantes con dataset 
  data_plus_LC<- cbind(theData,lattentclass=theClasses)
  #get the number of classes
  num_class= length(unique(theClasses))
  #get the number of features of the dataset (not includinf the latent class feature)
  num_feat<- ncol(theData)
  #get the mean values of each variable per each latent class  [1:14]
  mean_data_per_LC <- aggregate(data_plus_LC[,1:num_feat],
                                     list(data_plus_LC$lattentclass),
                                     mean)
  #eliminate the latent class feature
  rownames(mean_data_per_LC) <- mean_data_per_LC[,1]
  mean_data_per_LC <- mean_data_per_LC[,-1]
  #normalize
  norm_mean_data_per_LC <- scale(mean_data_per_LC)
  norm_mean_data_per_LC <- as.data.frame(norm_mean_data_per_LC)
  #get the min and max per feature 
  colMax <-  sapply(norm_mean_data_per_LC, max, na.rm = TRUE)
  colMin <- sapply(norm_mean_data_per_LC, min, na.rm = TRUE)
  #merge with max and min since radar needs it to plot
  data_for_radar <- rbind(colMax,colMin,norm_mean_data_per_LC)
  
  # Prepare color
  colors_border=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=1)
  colors_in=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=0.3)
  mytitle <- paste0("Cluster [", seq(1:num_class), "]")
  # Loop for each plot
  for(i in 1:num_class){
  
    # Custom the radarChart !
    radarchart( data_for_radar[c(1,2,i+2),], axistype=1, 
    
      #custom polygon
      pcol=colors_border[i] , pfcol=colors_in[i] , plwd=4, plty=1 , 
    
      #custom the grid
      cglcol="grey", cglty=1, axislabcol="grey",  cglwd=0.8,
    
      #custom labels
      vlcex=0.8,
      
      #title
      title=mytitle[i]
      )
  }
  ################################TSNE
  #sigo buscando como buscar unos buenos hyperparametros para el tsne
  #https://towardsdatascience.com/how-to-tune-hyperparameters-of-tsne-7c0596a18868
  tsne <- Rtsne(data_plus_LC[,-c(ncol(data_plus_LC)-1,ncol(data_plus_LC))], 
                dims = 2,
                perplexity=150,
                verbose=TRUE,
                max_iter = 2500,
                check_duplicates = FALSE)

  plot(tsne$Y, t='n', main="tsne")
  text(tsne$Y, labels=as.factor(data_plus_LC$outcome), col=colors[data_plus_LC$lattentclass])
  
}


```


# variaciones de HLCM (falta agregar el radar tsne)
```{r}
HLCM_LASSO_1SE <- HLCM(outcome ~.,trainingset, method=LASSO_1SE,
                       repetitions = 75,
                    family = "binomial")

LC_HLCM_LASSO_1SE <- getLattentClasses(HLCM_LASSO_1SE,testingset) 


HLCM_LASSO_1SE_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_LASSO_1SE,testingset)),
                                   "HLCM_LASSO_1SE",
                                   cex=0.8)

HLCM_EM_LASSO_1SE <- HLCM_EM(outcome ~.,trainingset, method=LASSO_1SE,
                       repetitions = 75,
                    family = "binomial")

HLCM_EM_LASSO_1SE_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_EM_LASSO_1SE,testingset)),
                                   "HLCM_EM_LASSO_1SE",
                                   cex=0.8)

```
 

```{r}


HLCM_model <- HLCM(outcome ~.,trainingset)

HLCM_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_model,testingset)),
                                   "HLCM",
                                   cex=0.8)
HLCM_EM_model <- HLCM_EM(outcome ~.,trainingset)

HLCM_EM_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_EM_model,testingset)),
                                   "HLCM_EM",
                                   cex=0.8)
```
 

#RADARS AND TSNE

```{r}
HLCM_EM_LASSO.SVM <- HLCM_EM(outcome ~ .,trainingset,
              hysteresis = 0.1,
              method=LASSO_1SE,
              classMethod=e1071::svm,
              classModel.Control=list(probability = TRUE),
              family="binomial")

testResult <-  predict(HLCM_EM_LASSO.SVM,testingset)

cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "LASSO SVM")


LC_LASSO.SVM <- getLattentClasses(HLCM_EM_LASSO.SVM,testingset)
#cat("Lattent classes found and its observations:",table(lattentclass))
#table(LC_LASSO.SVM)
#cat("Lattent classes found vs Outcome :")
#table(lc,testSet$outcome)
#testTable <- table(testResult < 0.5,testingset$outcome==0)
#epi.tests(testTable)
set.seed(42)
radar_and_tsne_from_latentclass(testingset,LC_LASSO.SVM)

```






```{r}
HLCM_EM_BSWIMS.SVM <- HLCM_EM(outcome ~ .,trainingset,
              hysteresis = 0.1,
              classMethod=e1071::svm,
              classModel.Control=list(probability = TRUE),
              NumberofRepeats = -1)


LC_BSWIMS.SVM <- getLattentClasses(HLCM_EM_BSWIMS.SVM,testingset)
#cat("Lattent classes found and its observations:")
table(lc)
#cat("Lattent classes found vs Outcome :")
table(lc,testSet$outcome)
testTable <- table(testResult < 0.5,testingset$outcome==0)
epi.tests(testTable)

testResult <-  predict(HLCM_EM_BSWIMS.SVM,testingset)

cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "BSWiMS SVM")

radar_and_tsne_from_latentclass(testingset,LC_BSWIMS.SVM)  
```



#OTHER MODELS TO COMPARE

```{r}
RF <- randomForest(outcome ~ ., data = trainingset, importance = TRUE)
# Predicting on train set
FR_result <- predict(RF, testingset, type = "class")
# Checking classification accuracy
#table(FR_result, testingset$outcome)  

cStats <- predictionStats_binary(cbind(testingset$outcome,FR_result),plotname = "RF")

```

 