---
title: "covid_mexico_clusters"
author: "eider"
date: "5/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r }
library("FRESA.CAD")
library(epiR)
library(fastAdaboost)
library(fmsb)
library(colormap)
library(Rtsne)

repetitions = 300
trainFraction = 0.90

setwd("C:/Users/eider/Dropbox/AD experiments/Covid_mexico/")
```


```{r}
COVID19MEXICO <- read.csv("200512COVID19MEXICO.csv",na.strings="99",stringsAsFactors = FALSE)

```

# Cleaning the data
 se seleccionan las variables para crear los modelos
 se seleccionan solo los casos positivos de covid
 se seleccionan los que tienen fecha de defuncion valida o entraron a cuidados intensivos como outome =1 , los demas 0
 
```{r}
colnamesMexico <- colnames(COVID19MEXICO)

varnames <- c("SEXO","TIPO_PACIENTE","FECHA_INGRESO","FECHA_SINTOMAS","FECHA_DEF",
              "INTUBADO","NEUMONIA","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR",
              "HIPERTENSION","OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA",
              "TABAQUISMO","OTRO_CASO","RESULTADO","UCI")


colnamesMexicoComplete <- COVID19MEXICO[,varnames]
colnamesMexicoComplete <- colnamesMexicoComplete[complete.cases(colnamesMexicoComplete),]

varforDiagnosisCovid <- c("SEXO","FECHA_DEF","EDAD","EMBARAZO","DIABETES","EPOC","ASMA","INMUSUPR","HIPERTENSION","OTRA_COM","CARDIOVASCULAR","OBESIDAD","RENAL_CRONICA","TABAQUISMO","RESULTADO","UCI")

# selecting just the variables needed 
datosDiagnosticoAlta <- colnamesMexicoComplete[,varforDiagnosisCovid] 

# solo los que dieron positivo a covid
datosDiagnosticoAlta <- subset(datosDiagnosticoAlta, RESULTADO == 1)
save(datosDiagnosticoAlta,file="datosDiagnosticoAlta.RDATA")
datosDiagnosticoAlta$RESULTADO <- NULL
#pacientes difuntos O si entraton a cuidados intensivos (UCI)
datosDiagnosticoAlta$outcome <- 1*(datosDiagnosticoAlta$FECHA_DEF != "9999-99-99" | datosDiagnosticoAlta$UCI == 1)

datosDiagnosticoAlta$FECHA_DEF <- NULL
datosDiagnosticoAlta$UCI <- NULL

table(datosDiagnosticoAlta$outcome)

```
 
#spliting the dataset 
- deberia de balancear mejor el split de train y test? 
```{r}
set.seed(42)
caseSet <- subset(datosDiagnosticoAlta,outcome==1)
controlSet <- subset(datosDiagnosticoAlta,outcome==0)

# 70 % of the UCI or death cases
samplesCasetrain <- sample(nrow(caseSet),0.70*nrow(caseSet))
# subseting the 70% 
trainingCaseset <- caseSet[samplesCasetrain,]

# 500 control cases
samplesControltrain <- sample(nrow(controlSet),500)
#subseting 500 control cases
trainingControlset <- controlSet[samplesControltrain,]
#merge 70% of data with cases and 500 training scores
trainingset <- rbind(trainingCaseset,trainingControlset)
#######################TEST

# selecting the remaining 30% of UCI and Death
controltestset <- controlSet[-samplesControltrain,]
# merge the remaining 30 and the controlsetttest
testingset <- rbind(caseSet[-samplesCasetrain,],controltestset[sample(nrow(controltestset),1000),])

```
 
#HCLM + LASSO_1SE
```{r}
HLCM_LASSO_1SE <- HLCM(outcome ~.,trainingset, method=LASSO_1SE,
                       repetitions = 75,
                    family = "binomial")

HLCM_LASSO_1SE_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_LASSO_1SE,testingset)),
                                   "HLCM_LASSO_1SE",
                                   cex=0.8)

HLCM_EM_LASSO_1SE <- HLCM_EM(outcome ~.,trainingset, method=LASSO_1SE,
                       repetitions = 75,
                    family = "binomial")

HLCM_EM_LASSO_1SE_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_EM_LASSO_1SE,testingset)),
                                   "HLCM_EM_LASSO_1SE",
                                   cex=0.8)

```

#pure HLCM (BSWIMS) 
```{r}


HLCM_model <- HLCM(outcome ~.,trainingset)

HLCM_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_model,testingset)),
                                   "HLCM",
                                   cex=0.8)
HLCM_EM_model <- HLCM_EM(outcome ~.,trainingset)

HLCM_EM_bs <- predictionStats_binary(cbind(testingset$outcome,
                                   predict(HLCM_EM_model,testingset)),
                                   "HLCM_EM",
                                   cex=0.8)
```
 
 ########### RANDOMCV
 
```{r}
repetitions = 100
trainFraction = 0.70
#lasso 1se
HLCM_EM_cv <- randomCV(datosDiagnosticoAlta,
                       "outcome",
                       fittingFunction=HLCM_EM,
                       trainFraction = trainFraction,
                       repetitions = repetitions)
```


```{r}
HLCM_EM_LASSO.SVM <- HLCM_EM(outcome ~ .,trainingset,
              hysteresis = 0.1,
              method=LASSO_1SE,
              classMethod=e1071::svm,
              classModel.Control=list(probability = TRUE),
              family="binomial")

testResult <-  predict(HLCM_EM_LASSO.SVM,testingset)
lattentclass <- apply(attr(testResult,"probabilities"),1,which.max)
table(lattentclass)
table(lattentclass,testingset$outcome)
testTable <- table(testResult < 0.5,testingset$outcome==0)
epi.tests(testTable)
cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "LASSO SVM")

getLattentClasses <- function(lattentClassModel,testSet){
  #get the results from test set
  testResult <-  predict(lattentClassModel,testSet)
  #from each observation get the max probabilty to belong to a given LC
  lattentclass <- apply(attr(testResult,"probabilities"),1,which.max)
  cat("Lattent Classes found: ",length(unique(lattentclass)))
  lattentclass
  
}
  
lc <- getLattentClasses(HLCM_EM_LASSO.SVM,testingset)
cat("Lattent classes found and its observations:",table(lattentclass))
table(lc)
cat("Lattent classes found vs Outcome :")
table(lc,testSet$outcome)
testTable <- table(testResult < 0.5,testingset$outcome==0)
epi.tests(testTable)

```



#¿Puedes comparar las dos: LASSO vs BSWiMS?
RADAR
preguntas:
deberia de filtrar a los clusters que tengan mayor a algun treshold de observaciones?
alguna recomendacion para el hyperparamater del TSNE?
```{r}

#combinamos las clases resultantes del testset con 
data_plus_LC<- cbind(testingset,lattentclass=lattentclass)
#get the number of classes
num_class= length(unique(lattentclass))
#get the mean values of each cluster for each variable [1:14]
mean_data_for_radar_1 <- aggregate(data_plus_LC[,1:14],
                                   list(data_plus_LC$lattentclass),mean)
#eliminate the latent class feature
rownames(mean_data_for_radar_1) <- mean_data_for_radar_1[,1]
mean_data_for_radar_1 <- mean_data_for_radar_1[,-1]
#normalize
mean_data_for_radar_1 <- scale(mean_data_for_radar_1)
mean_data_for_radar_1 <- as.data.frame(mean_data_for_radar_1)
#get the min and max per feature 
colMax <-  sapply(mean_data_for_radar_1, max, na.rm = TRUE)
colMin <- sapply(mean_data_for_radar_1, min, na.rm = TRUE)

#merge
mean_data_for_radar_1 <- rbind(colMax,colMin,mean_data_for_radar_1)


# Prepare color
colors_border=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=1)
colors_in=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=0.3)

# Prepare title
mytitle <- c("cluster 1","cluster 2","cluster 3","cluster 4")

# Loop for each plot
for(i in 1:num_class){

  # Custom the radarChart !
  radarchart( mean_data_for_radar_1[c(1,2,i+2),], axistype=1, 
  
    #custom polygon
    pcol=colors_border[i] , pfcol=colors_in[i] , plwd=4, plty=1 , 
  
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey",  cglwd=0.8,
  
    #custom labels
    vlcex=0.8,
    
    #title
    title=mytitle[i]
    )
}
#for ploting
colors = colormap(colormap=colormaps$viridis, nshades=num_class, alpha=1)

tsne <- Rtsne(data_plus_LC[,-15], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500,check_duplicates = FALSE)

plot(tsne$Y, t='n', main="tsne")
text(tsne$Y, labels=data_plus_LC$outcome, col=colors[data_plus_LC$lattentclass])
```
 


```{r}

#data must be in feature-outcome format
radar_and_tsne_from_latentclass <- function(theData,theClasses){
  #combinamos las clases resultantes con dataset 
  data_plus_LC<- cbind(theData,lattentclass=theClasses)
  #get the number of classes
  num_class= length(unique(theClasses))
  #get the number of features of the dataset (not includinf the latent class feature)
  num_feat<- ncol(theData)
  #get the mean values of each variable per each latent class  [1:14]
  mean_data_per_LC <- aggregate(data_plus_LC[,1:num_feat],
                                     list(data_plus_LC$lattentclass),
                                     mean)
  #eliminate the latent class feature
  rownames(mean_data_per_LC) <- mean_data_per_LC[,1]
  mean_data_per_LC <- mean_data_per_LC[,-1]
  #normalize
  norm_mean_data_per_LC <- scale(mean_data_per_LC)
  norm_mean_data_per_LC <- as.data.frame(norm_mean_data_per_LC)
  #get the min and max per feature 
  colMax <-  sapply(norm_mean_data_per_LC, max, na.rm = TRUE)
  colMin <- sapply(norm_mean_data_per_LC, min, na.rm = TRUE)
  #merge with max and min since radar needs it to plot
  data_for_radar <- rbind(colMax,colMin,norm_mean_data_per_LC)
  
  # Prepare color
  colors_border=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=1)
  colors_in=colormap(colormap=colormaps$viridis, nshades=num_class, alpha=0.3)
  mytitle <- paste0("Cluster [", seq(1:num_class), "]")
  # Loop for each plot
  for(i in 1:num_class){
  
    # Custom the radarChart !
    radarchart( data_for_radar[c(1,2,i+2),], axistype=1, 
    
      #custom polygon
      pcol=colors_border[i] , pfcol=colors_in[i] , plwd=4, plty=1 , 
    
      #custom the grid
      cglcol="grey", cglty=1, axislabcol="grey",  cglwd=0.8,
    
      #custom labels
      vlcex=0.8,
      
      #title
      title=mytitle[i]
      )
  }
  ################################TSNE
  #sigo buscando como buscar unos buenos hyperparametros para el tsne
  #https://towardsdatascience.com/how-to-tune-hyperparameters-of-tsne-7c0596a18868
  tsne <- Rtsne(data_plus_LC[,-c(ncol(data_plus_LC)-1,ncol(data_plus_LC))], 
                dims = 2,
                perplexity=150,
                verbose=TRUE,
                max_iter = 2500,
                check_duplicates = FALSE)

  plot(tsne$Y, t='n', main="tsne")
  text(tsne$Y, labels=as.factor(data_plus_LC$outcome), col=colors[data_plus_LC$lattentclass])
  
}
set.seed(42)
radar_and_tsne_from_latentclass(testingset,lattentclass)

```


```{r}
HLCM_EM_BSWIMS.SVM <- HLCM_EM(outcome ~ .,trainingset,
              hysteresis = 0.1,
              classMethod=e1071::svm,
              classModel.Control=list(probability = TRUE),
              NumberofRepeats = -1)

HLCM_EM_BSWIMS.SVM$selectedfeatures
###########train
trainResult <-  predict(HLCM_EM_LASSO,trainingset)
#get the probabilites given a latent class 
train_probs <- head(attr(trainResult,"probabilities"))
train_probs
#get the cluster given the max probabilities
apply(train_probs,1,which.max)

##get the probabilites given a latent class 
lattentclass <- apply(attr(trainResult,"probabilities"),1,which.max)
table(lattentclass)
#compare the each cluster distribution of outcome 
table(lattentclass,trainingset$outcome)
#confusion matrix of trainset
testTable <- table(trainResult < 0.5,trainingset$outcome==0)
epi.tests(testTable)


#############TEST SET
testResult <-  predict(HLCM_EM_LASSO,testingset)
lattentclass <- apply(attr(testResult,"probabilities"),1,which.max)
table(lattentclass)
table(lattentclass,testingset$outcome)
testTable <- table(testResult < 0.5,testingset$outcome==0)
epi.tests(testTable)
cStats <- predictionStats_binary(cbind(testingset$outcome,testResult),plotname = "BSWiMS SVM")
##################nuevo
lc <- getLattentClasses(HLCM_EM_LASSO,testingset)
cat("Lattent classes found and its observations:",table(lattentclass))
table(lc)
cat("Lattent classes found vs Outcome :")
table(lc,testSet$outcome)
testTable <- table(testResult < 0.5,testingset$outcome==0)
epi.tests(testTable)


```

 